SMT-Based Generation of Symbolic Automata \thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2019-03-27 2017-Xudong-SMT/ActaInformatica.tex
Sketch of the proofs

\begin{proof}
  Let $pn=<...>$ be a pNet. The set of states 
  managed by the two algorithms both are subsets of the cartesian
  product of the pLTS states, as defined in \pageref{def-states}. In a
  similar way, the set of state variables are also the same. Comparing
  transitions requires a little more care, because within the
  MakeTransitions function, fresh variables are created everytime a
  synchronisation vector is used (this occurs within the $MkPred$ and
  $matching$ functions). However, given a global state, the call of
  $MakeTransitions$ will always returns the same set of OTs, up to
  renaming of these fresh variables.

  Let us define the following sets of OTs, as computed by the 2
  algorithms:
  \begin{itemize}
    \item $L$ = original set of transitions by Algorithm 1, before pruning
      (i.e. before line 11),
    \item $L1-reach$ = final set of OTs of Algorithm 1, as computed
        by line 12.
      \item $L2-reach$ = final set of OTs of Algorithm 2, as computed
          by line 13.
  \end{itemize}

@@

  Now remark that without assignments,
  Theorem~\ref{Th:CorrectnessWithoutAssignments} boils to proving

Let $L$, $L1$, $L_{Smart}$ respectively mean the result open transition set of original algorithm before pruning, original algorithm after pruning and smart algorithm. Similarly, $State(L)$, $State(L1)$, $State(L_{Smart})$ mean the result global state set of them. According to the algorithm, there has $L_{Smart}\in L$ and $State(L_{Smart})\in State(L)$.

Suppose that $f$: $State(L1)\rightarrow State(L_{Smart})$.
\begin{enumerate}
\setlength{\itemindent}{3.5em}
	\item[(Surjective)] If $\ostate{\tilde{s_0}}=Initial(p)$ then $f(\ostate{\tilde{s}})=\ostate{\tilde{t_0}}=Initial(p)$ must exists, the transitions started from them fulfill that
	\[\forall \ostate{\tilde{t_0'}}, \ostate{\tilde{t_0}}\xrightarrow{v_j}\ostate{\tilde{t_0'}}\in L_{Smart},\quad \exists\ostate{\tilde{s'}},\ostate{\tilde{s_0}}\xrightarrow{v_i}\ostate{\tilde{s_0'}}\in L.\]
	The state $\ostate{\tilde{t'}}\in State(L_{Smart})$ and as $Pred_j$ of $\ostate{\tilde{t}}\xrightarrow{v_j}\ostate{\tilde{t_0'}}$ is satisfiable, then if the $Pred_i$ of $\ostate{\tilde{s_0}}\xrightarrow{v_i}\ostate{\tilde{s_0'}}$ is also satisfiable, the state $\ostate{\tilde{s_0'}}$ is reachable in the original open automaton s.t. $\ostate{\tilde{s_0'}}\in State(L1)$, we can construct the mapping
	\[f(\ostate{\tilde{s_0'}})=\ostate{\tilde{t_0'}}.\]
	If the $Pred_i$ of $\ostate{\tilde{s_0}}\xrightarrow{v_i}\ostate{\tilde{s_0'}}$ is not satisfiable then two transitions are not matched and $\ostate{\tilde{s_0'}}\notin State(L1)$.

@@

%Intuitively, $\forall \ostate{\tilde{s}}\in\textit{Explored}=State(L)\text{, there must exist a transition}\ostate{\tilde{s_i}}\xrightarrow{v_i}\ostate{\tilde{s}}\text{ and }\forall \ostate{\tilde{s}}\in\textit{Explored}_{Smart}=State(L_{Smart})\text{, }\exists\ostate{\tilde{pre(s)}}\text{, }\ostate{\tilde{pre(s)}}\xrightarrow{v}\ostate{\tilde{s}}\text{ where \textit{Pred} is satisfiable and the state}\ostate{\tilde{pre(s)}}\text{ has two cases:}$
%\begin{enumerate}
%	\item[(1)] $\ostate{\tilde{pre(s)}}=\ostate{\tilde{s}}$
%	\item[(2)] $\ostate{\tilde{pre(s)}}\in\textit{Explored}\text{-}\{\ostate{\tilde{s}}\}$
%\end{enumerate}
%By induction, there must exists a sequence $\ostate{\tilde{s_0}}\xrightarrow{v_i}\ostate{\tilde{s}}$ where all the $\textit{Pred}_i$ are satisfiable.
%
%Then $\forall\ostate{\tilde{t}}\textit{Explored}=State(L)$, there are also two cases:
%\begin{enumerate}
%	\item[(1)] $\ostate{\tilde{t}}\in State(L1)$
%	
%	By definition of reachability, there exists a sequence $\ostate{\tilde{s_0}}...\xrightarrow{v_j}...\ostate{\tilde{t}}$ where all the $\textit{Pred}_j$ are satisfiable.
%	
%	$\Rightarrow \ostate{\tilde{t}}\in State(L_{Smart})$
%	\item[(2)] $\ostate{\tilde{t}}\in State(L)\text{-}State(L1)$, then $\forall \ostate{\tilde{u}}\in\textit{Explored}-\{\ostate{\tilde{t}}\}$,
%	\begin{enumerate}
%		\item[(1)] $\ostate{\tilde{u}}\in State(L)\text{-}State(L1)$.
%		\item[(2)] $\ostate{\tilde{u}}\in State(L1)$, but there is no transition $\ostate{\tilde{u}}\xrightarrow{v_u}\ostate{\tilde{t}}$.
%		\item[(3)] $\ostate{\tilde{u}}\in State(L1)$, there exists transition $\ostate{\tilde{u}}\xrightarrow{v_u}\ostate{\tilde{t}}$, but the $\textit{Pred}_u$ is unsatisfiable.
%	\end{enumerate}
%	By induction, there is no sequence from $\ostate{\tilde{s_0}}$ to $\ostate{\tilde{t}}$.
%	
%	$\Rightarrow \ostate{\tilde{t}}\notin State(L_{Smart})$
%\end{enumerate}
%Overall, $State(L1)$ is equal to $State(L_{Smart})$.
\end{proof}

@@

\TODO{Xudong: proof - to be checked}

\begin{proof}
In Theorem 2, we have already proved the result set of global states of two algorithm are the same without the assignments.

>>>>

\ERIC{Sketch of the proof:}

\begin{proof}
  Let $pn=<...>$ be a pNet. The set of states 
  managed by the two algorithms both are subsets of the cartesian
  product of the pLTS states, as defined in \pageref{def-states}. In a
  similar way, the set of state variables are also equal. Comparing
  transitions requires a little more care, because within the
  MakeTransitions function, fresh variables are created everytime a
  synchronisation vector is used (this occurs within the $MkPred$ and
  $matching$ functions). However, given a global state, the call of
  $MakeTransitions$ will always returns exactly the same set of OTs, up to
  renaming of these fresh variables\footnote{I think this does not
    need any specific proof, given that it is the same function,
    called with the same arguments}.

  Let us define the following sets of OTs, as computed by the 2
  algorithms:
  \ERIC{not sure whether all these sets are useful, but it's good to
    have them precisely defined for the moment}
  \begin{itemize}
    \item $L1$ = original set of transitions by Algorithm 1
      (i.e. before line 11), 
    \item $L1-sat$ = original set of transitions by Algorithm 1, after
      satisfiability check (i.e. after line 11),
    \item $L1-reach$ = final set of OTs of Algorithm 1, as computed
      by line 12.
    \item $L2-sat$ = original set of transitions by Algorithm 2
      (i.e. before line 13), 
    \item $L2-reach$ = final set of OTs of Algorithm 2, as computed
          by line 13.
  \end{itemize}

@@

  From the structure of the main loops of the algorithms, we know that
  $L1-sat \subset L2-sat$ and subsequently $States(L1-sat) \subset       
  States(L2-sat)$.
  \qed
\end{proof}

    \medskip
  \noteInEM{I keep below the original proof draft from Xudong, though
    I'm not sure if parts of it can be reused}
\begin{proof}
%% Let $L$, $L1$, $L_{Smart}$ respectively mean the result open transition set of original algorithm before pruning, original algorithm after pruning and smart algorithm. Similarly, $State(L)$, $State(L1)$, $State(L_{Smart})$ mean the result global state set of them. According to the algorithm, there has $L_{Smart}\in L$ and $State(L_{Smart})\in State(L)$.

We will construct a mapping $f$: $State(L1-reach)\rightarrow
State(L2-reach)$.
\ERIC{I do not understand this, as states are identical. You would
  rather need mapping betwen transitions ?}
\begin{enumerate}
\setlength{\itemindent}{3.5em}
	\item[(Surjective)] If $\ostate{\tilde{s_0}}=Initial(p)$ then $f(\ostate{\tilde{s}})=\ostate{\tilde{t_0}}=Initial(p)$ must exists, the transitions started from them fulfill that
	\[\forall \ostate{\tilde{t_0'}}, \ostate{\tilde{t_0}}\xrightarrow{v_j}\ostate{\tilde{t_0'}}\in L_{Smart},\quad \exists\ostate{\tilde{s'}},\ostate{\tilde{s_0}}\xrightarrow{v_i}\ostate{\tilde{s_0'}}\in L.\]
	The state $\ostate{\tilde{t'}}\in State(L_{Smart})$ and as $Pred_j$ of $\ostate{\tilde{t}}\xrightarrow{v_j}\ostate{\tilde{t_0'}}$ is satisfiable, then if the $Pred_i$ of $\ostate{\tilde{s_0}}\xrightarrow{v_i}\ostate{\tilde{s_0'}}$ is also satisfiable, the state $\ostate{\tilde{s_0'}}$ is reachable in the original open automaton s.t. $\ostate{\tilde{s_0'}}\in State(L1)$, we can construct the mapping
	\[f(\ostate{\tilde{s_0'}})=\ostate{\tilde{t_0'}}.\]
	If the $Pred_i$ of $\ostate{\tilde{s_0}}\xrightarrow{v_i}\ostate{\tilde{s_0'}}$ is not satisfiable then two transitions are not matched and $\ostate{\tilde{s_0'}}\notin State(L1)$.

@@

\end{proof}

@@

\TODO{Formalize the proof}

\ERIC{Proof sketch:}

  \begin{itemize}
   \item $L1$ = original set of transitions by Algorithm 1, after
      satisfiability check (i.e. after line 11), using assignments
      from the whole state space.
   \item $L2$ = original set of transitions by Algorithm 1, after
      satisfiability check (i.e. after line 11), using assignments
      from the original on-the-fly search.

    \item $L1-reach$ = final set of OTs of Algorithm 1, as computed
        by an iterative application of makeReachableSubAutomaton (line
        12) up to fixpoint.
      \item $L2-reach$ = final set of OTs of Algorithm 2, as computed
          by line 13, as computed
        by an iterative application of makeReachableSubAutomaton (line
        16) up to fixpoint.
  \end{itemize}

  Proof similar to the previous case still apply here, but adapted as
  we are building a different starting point :

  - Define the function $States$ as before, and  $Assignments : OT^* \# State -> $ the function
  computing the set of assignments of a global state from a given set of OTs.

  - The structure implicitely computed by the algorithm (comprised in
  the structure of an Open Automaton) is a triple $ Aut= <States, OTs,
  Assigns>$.

  - when building $L1$ and $L2$ the 2 algorithms compute the initial
  $Aut_0$ structures, without using previous assignments, as
  explained in section~\ref{section:implementation}.

  - The function $Refine : Aut->Aut$, implemented as
  $makeReachableSubAutomaton(S0,OTs)$, uses the assignments of the
  previous step to refine the computed reachable space, reducing all 3
  components of the triple: by pruning some Unsat OTs, it potentialy
  removes the corresponding assignments, and their target states.

  - the whole computation takes place within a finite domain: Global
  states are within the cartesian product of the pLTS states; OTs are
  always subsets of $L1$, and Assignments are a fixed finite set of
  symbolic assignements of expressions to state variables,
  syntacticaly contained in the OTs in $L1$; so the fixpoint
  computation terminates.

  - the result $Aut$ contains all reachable OTs and States from $L1$,
  and only these (proof should be similar to that of Th2).

  \medskip
  \noteInEM{I keep below the original proof draft from Xudong, though
    think it does not apply any more}
\begin{proof}
In Theorem 2, we have already proved the result set of global states of two algorithm are the same without the assignments.
